\documentclass[11pt]{article}

\input{../common.tex}
\begin{document}
\customheading{\today}{Proof by contradiction: Rationality of Numbers}{Fall 2024}{CS30 - Discrete Maths}

\section*{}
\begin{enumerate}
    \item  Prove that $log_{2}3$ is irrational.
    \begin{proof}
        Suppose $log_{2}3$ were rational, i.e. $log_{2}3 = \frac{p}{3}$ where $p$ and $q$ are some positive integers. Then $3 = 2^{\frac{p}{q}} \equiv 3^{q} = 2^{p}$. An odd number cannot be equal to an even number hence $log_{2}3$ is irrational. \\

        Since $log_{2}3$ is an irrational number, $2log_{2}3$ is irrational.
        But it is not always the case. For example, let $x = \sqrt{2}$ and $y=log_{2}3$ then
        \begin{equation}
            x^{y} = (\sqrt{2})^{2log_{2}3} = \sqrt{2^{2}}^{log_{2}3} = 2^{log_{2}3} = 3
        \end{equation}
    \end{proof}

    \item Prove that $\sqrt{6}$ is irrational.
    \begin{proof}
        Suppose $\sqrt{6}$ were rational, i.e. $\sqrt{6} = \frac{p}{q}$ where $p$ and $q$ are some positive integers, and at least one of them is odd.If you square all sides, you will have, $ 6 = (\frac{p}{q})^{2} \equiv 6q^{2} = p^{2} = 2\cdot3q^{2} = p^{2}$. Therefore $p^{2}$ must be even, then $p$ is even. Let's say $p=2k \Rightarrow p^{2} = 4k^{2} = 2\cdot 3q^{2} \equiv \frac{2}{3}k^{2} = q^{2}$, but $q^{2}$ is a multiple of $2$ hence $q^{2}$ is even, and so is $q$. This is a contradtion because we assumed that either $p$ or $q$ is odd.
    \end{proof}

    \item Let $a$ and $b$ be two positive integers such that $a > b$, and $\sqrt{a}$ and $\sqrt{b}$ are both irrational numbers. Show that $\sqrt{a} + \sqrt{b}$ is always irrational.
    \begin{proof}
        Suppose $\sqrt{a} + \sqrt{b} = q$ Consider the product of difference of squares
        \begin{gather}
            q \cdot (\sqrt{a} - \sqrt{b}) = a-b
        \end{gather}
        Therefore, if $q$ is rational, then $r:=(\sqrt{a} - \sqrt{b}) = \frac{a-b}{q}$ is also rational.
    \end{proof}
    \begin{proof}
        Another style of proof:
        Suppose $\sqrt{a} + \sqrt{b} = q$ is rational
        \begin{gather}
            \sqrt{a} = q-\sqrt{b} \\
            a = (q - \sqrt{b})^{2} \\
            a = q^{2} - 2q\sqrt{b} + b \\
            \sqrt{b} = \frac{b-a+q^{2}}{2q}
        \end{gather}

        Since $a$ and $b$ are integers, the result of RHS is an integer, and $\sqrt{b}$ equals an integer divided by a rational number, hence $\sqrt{b}$ is a rational number, however, we initially assumed that $b$ is rational hence the contradiction.
    \end{proof}
\end{enumerate}

\end{document}